---
title: "Audit Final"
author: "Abhinav Ram Bhatta"
date: "2023-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#Loading data from all files and storing in variables for easy access

library(readr)
folder_path <- "/Users/abhinavram/Documents/Sem 2 UIC/IDS 523 Audit/auditanalytics_final_files"

csv_files <- list.files(folder_path, pattern = "\\.csv$")

for (file in csv_files) {
    
  # Removing the file extension and creating a variable name
  var_name <- tools::file_path_sans_ext(file)
  
  # Reading the CSV file and assigning it to a variable with the generated name
  assign(var_name, read_csv(file.path(folder_path, file)), envir = .GlobalEnv)
}

```


```{r}
library(plotluck)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(gridExtra)
library(lubridate)
```



# Question 1: Preliminary Tests of Transactions Controls

```{r}


# Creating a data frame with the inv_extended column
YE_Inventory_df <- perpetual_inventory_ledger %>% group_by(sku) %>% inner_join(sales_journal, by = "sku") %>% slice(n()) %>% mutate(inv_extended = unit_cost * stock_on_hand) %>% ungroup() %>% select(inv_extended)

# Calculating the sum of inv_extended
YE_Inventory <- sum(YE_Inventory_df$inv_extended)

Sales <- sum(sales_journal$collection_amount)
Sales_Returns <- sum(sales_journal$sales_return * sales_journal$sales_unit_price * sales_journal$sales_count)
AR <- sum(fyear_end_ar_ledger$amount)

Begin_Inventory <- fyear_begin_inventory_ledger %>% group_by(sku) %>% mutate(inv_extended = unit_cost * stock_on_hand) %>% ungroup() %>% select(inv_extended) %>% sum()
Cash <- sales_journal %>% filter(cash_not_ar == 1) %>% select(collection_amount) %>% sum()

# Sales
p1 <- plotluck(sales_journal, collection_amount~1) +  ggtitle("Plot of Sales") + xlab("Sales amount") + ylab("Density")

# Cash
p2 <- plotluck(deposit_daily,deposit_amount~1) +  ggtitle("Plot of Cash") +xlab("Cash Amount") + ylab("Density")

# AR
p3 <- plotluck(fyear_end_ar_ledger,amount~1) + ggtitle("Plot of Accounts Receivable") + xlab("Amount") + ylab("Density")

# Inventory
p4 <- plotluck(YE_Inventory_df, inv_extended~1) + ggtitle("Plot of Inventory")

grid.arrange(
  arrangeGrob(p1, p2, ncol = 2),
  arrangeGrob(p3, p4, ncol = 2),
  nrow = 2
)

# Returns
p5 <- plot(density(sales_journal$sales_return * sales_journal$sales_unit_price), main = "Returns")

# COGS
p6 <- plot(density(sales_journal$sales_count * sales_journal$unit_cost), main = "COGS")

```


# Question 2: Sales Duplicates and Ommissions Error Rates

```{r}

#Invoice numbers
dup_invoice <-  sales_journal[duplicated(sales_journal$invoice_no), ] 
n <- nrow(dup_invoice)
dup_invoice_per <- (n / nrow(sales_journal))*100
cat("\n % of duplicate Invoice number = ", dup_invoice_per)

#Shipping numbers
dup_shipping <-  shipments_journal[duplicated(shipments_journal$shipper_no), ] 
n <- nrow(dup_shipping)
dup_shipping_per <- (n / nrow(shipments_journal))*100
cat("\n % of duplicate Shipping numbers = ", dup_shipping_per)


#Collection Receipt numbers
dup_collection <-  collections_journal[duplicated(collections_journal$collection_no), ] 
n <- nrow(dup_collection)
dup_collection_per <- (n / nrow(collections_journal))*100
cat("\n % of duplicate Collection receipt numbers = ", dup_collection_per)


#Customers with credit balances
dup_credit <-  fyear_end_ar_ledger[duplicated(fyear_end_ar_ledger$invoice_no), ] 
n <- nrow(dup_credit )
dup_credit_per <- (n / nrow(fyear_end_ar_ledger))*100
cat("\n % of duplicate Customers with credit balances = ", dup_credit_per )
```




```{r}
#Invoice numbers
invoice <- as.numeric(substring(sales_journal$invoice_no, 2)) 
invoice_min <- as.numeric(min(invoice))
invoice_max <- as.numeric(max(invoice))

omit <- as.data.frame(setdiff(invoice_min:invoice_max, invoice)) 
n <- nrow(omit)
omi_invoice_per <- (n / nrow(sales_journal))*100
cat("\n % of omitted Invoice numbers = ", omi_invoice_per)

#Shipping numbers
shipping <- as.numeric(substring(shipments_journal$shipper_no, 2)) 
shipping_min <- as.numeric(min(shipping))
shipping_max <- as.numeric(max(shipping))

omit <- as.data.frame(setdiff(shipping_min:shipping_max, shipping)) 
n <- nrow(omit)
omi_shipping_per <- (n / nrow(shipments_journal))*100
cat("\n % of omitted Shipping numbers = ", omi_shipping_per)


#Collection Receipt numbers
collection <- as.numeric(substring(collections_journal$collection_no, 2)) 
collection_min <- as.numeric(min(collection))
collection_max <- as.numeric(max(collection))

omit <- as.data.frame(setdiff(collection_min:collection_max, collection)) 
n <- nrow(omit)

omi_collection_per <- (n / nrow(collections_journal))*100
cat("\n % of omitted Collection receipt numbers = ", omi_collection_per)


#Customers with credit balances 
credit <- as.numeric(substring(fyear_end_ar_ledger$invoice_no, 2)) 
credit_min <- as.numeric(min(credit))
credit_max <- as.numeric(max(credit))

omit <- as.data.frame(setdiff(credit_min:credit_max, credit)) 
n <- nrow(omit)

omi_credit_per <- (n / nrow(fyear_end_ar_ledger))*100
cat("\n % of omitted Customers with credit balances  numbers = ", omi_credit_per)
```




# Question 3: Error Rates in Sales Amounts

```{r}

MTE<-(100000/sum(sales_journal$sales_count*sales_journal$sales_unit_price))
MTE #1.20% is the intolerable error
confidence <- .95
n <- (log(1-confidence))/log(1-MTE)
n<-round(n)
sample_size<-n


rw_credit_sales <- read.csv("/Users/abhinavram/Documents/Sem 2 UIC/IDS 523 Audit/auditanalytics_final_files/real_world_credit_sales.csv",na.strings="0",stringsAsFactors=FALSE) %>% select(invoice_no,sales_unit_price,sales_count)

rw_cash_sales <- read.csv("/Users/abhinavram/Documents/Sem 2 UIC/IDS 523 Audit/auditanalytics_final_files/real_world_cash_sales.csv",na.strings="0",stringsAsFactors=FALSE) %>% select(invoice_no,sales_unit_price,sales_count)

#Joining real world Credit Sales and Cash Sales
rw_join_sales <- rbind(rw_credit_sales,rw_cash_sales) 


#Discovery sample for Sales Journal
discovery_sample_sales <- sales_journal[sample(nrow(sales_journal), sample_size),] %>%
select(invoice_no,sales_unit_price,sales_count)

#Discovery sample for Real world Credit & Sash files combined
discovery_sample_rw_sales <- rw_join_sales[sample(nrow(rw_join_sales), sample_size),] %>%
select(invoice_no,sales_unit_price,sales_count)

#Inner joining real world credit sales and real world cash sales
audited_sample <-left_join(discovery_sample_sales,discovery_sample_rw_sales,by="invoice_no") %>%
select(invoice_no,sales_unit_price.x,sales_count.x,sales_unit_price.y,sales_count.y)

exceptions <-audited_sample %>%filter(sales_unit_price.x != sales_unit_price.y |
sales_count.x != sales_count.y
)
error_rate_sales <- nrow(exceptions) / nrow(audited_sample)
cat("\n Discovery: error rate in sales:", error_rate_sales)

error_sales_amt <-
100 * sum(exceptions$sales_count.y * exceptions$sales_unit_price.y) /
sum(audited_sample$sales_unit_price.x * audited_sample$sales_count.x)
cat("\n Discovery: amount for sales sample is overstated (i.e. is in error):",
prettyNum(error_sales_amt, big.mark=","),
" %")
```



# Question 4: Compute dollar and percentage totals for:

```{r}
#Converting to date
sales_journal$invoice_date <- as.Date(sales_journal$invoice_date)
sales_journal$collection_date <- as.Date(sales_journal$collection_date)

#Calculating days_outstanding
sales_journal$days_outstanding <- as.numeric(sales_journal$collection_date - sales_journal$invoice_date)

#Calculating dollar totals 
ARUnder30  <- sum(sales_journal$collection_amount[which(sales_journal$days_outstanding <  30)])
ARBet30n60 <- sum(sales_journal$collection_amount[which(sales_journal$days_outstanding >= 30 & sales_journal$days_outstanding <= 60)])
ARAbove60  <- sum(sales_journal$collection_amount[which(sales_journal$days_outstanding >  60)])

#calculating percentage totals
ARUnder30per  <- (ARUnder30  / sum(sales_journal$collection_amount)) *100
ARBet30n60per <- (ARBet30n60 / sum(sales_journal$collection_amount)) *100
ARAbove60per  <- (ARAbove60  / sum(sales_journal$collection_amount)) *100

#List of invoices over 60 
ARAbove60in <- sales_journal$invoice_no[which(sales_journal$days_outstanding >  60)]
as.data.frame(ARAbove60in)

ARAbove60in <- subset(sales_journal, days_outstanding >  60)


#Printing dollar and percentage totals
cat("a) All Invoices under 30 days old:\n")
cat("Dollar Total: ", ARUnder30, "\n")
cat("Percentage Total: ", ARUnder30per, "%\n\n")

cat("b) All Invoices between 30 to 60 days old:\n")
cat("Dollar Total: ", ARBet30n60, "\n")
cat("Percentage Total: ", ARBet30n60per, "%\n\n")

cat("c) All Invoices over 60 days old:\n")
cat("Dollar Total: ", ARAbove60, "\n")
cat("Percentage Total: ", ARAbove60per, "%\n\n")


# Filtering rows with missing or non-alphanumeric customer_no values
ARAbove60in <- ARAbove60in[!is.na(ARAbove60in$customer_no), ]

#Count of frequency of each customer_no
customer_count <- table(ARAbove60in$customer_no)

#Creating a bar plot
barplot(customer_count, main="Customer Numbers for Invoices over 60 Days",
        xlab="Customer Number", ylab="Frequency", las=2, cex.names=0.8)

plot(ARAbove60in$days_outstanding)

plot(seq_along(ARAbove60in$invoice_no), ARAbove60in$days_outstanding,
      main="Invoice over 60 days", 
      xlab="Number of Invoices", ylab="Number of Days")

hist(as.numeric(ARAbove60in$days_outstanding), main="Invoice over 60 days", 
      xlab="Number of Invoices", ylab="Number of Days"
)
```




# Question 5: Identify all customer balances greater than their credit limit.  

```{r}
#Calculating credit limits
cr_limit <- sales_journal %>%
  group_by(customer_no) %>%
  summarize(ann_sales = sum(sales_extended),
            time_to_pay = mean(as.numeric(collection_date - invoice_date), na.rm = TRUE),
            credit_limit = 10 * round(ann_sales / time_to_pay, digits = -2))

#Calculating customer AR balance
customer_age_ar <- sales_journal %>%
  group_by(customer_no) %>%
  summarize(total = sum(collection_amount))

#Joining AR balance with credit limits and checking for customers over the limit
OverlimitTable <- full_join(cr_limit, customer_age_ar, by = "customer_no") %>%
  mutate(OverLimit = ifelse((total > credit_limit), "Yes", "No"))

#Identifying customers with balances greater than their credit limit
problematic_customers <- OverlimitTable %>%
  filter(OverLimit == "Yes")

#Displaying problematic customers
cat("Customers with balances greater than their credit limit:\n")
print(problematic_customers)

#Checking if any customers are substantially over their credit limit
substantially_over_limit <- problematic_customers %>%
  mutate(over_limit_percentage = (total - credit_limit) / credit_limit) %>%
  filter(over_limit_percentage > 0.5)
#Customers with over 50% of their credit limit as a threshold

#Displaying substantially over limit customers
if (nrow(substantially_over_limit) > 0) {
  cat("\nCustomers substantially over their credit limit (over 50%):\n")
  print(substantially_over_limit)
} else {
  cat("\nNo customers are substantially over their credit limit (over 50%).\n")
}


```



# Question 6: Perform a sales cutoff test

```{r}
fisyear_begin_2021 <- as.Date("2021-01-01")
fisyear_end_2021 <- as.Date("2021-12-31")
fisyear_begin_2022 <- as.Date("2022-01-01")
fisyear_end_2022 <- as.Date("2022-12-31")

sales_journal$invoice_date <- as.Date(sales_journal$invoice_date)
sales_journal$shipper_date <- as.Date(sales_journal$shipper_date)

sales_cutoff_errors <- sales_journal %>% 
  filter(invoice_date <= fisyear_end_2021 & shipper_date >= fisyear_begin_2022) %>% 
  select(customer_no, invoice_no, invoice_date, shipper_no, shipper_date, sales_extended)

sales_cutoff_errors

cutoff_error_count <- nrow(sales_cutoff_errors)
cutoff_error_percentage <- nrow(sales_cutoff_errors) / nrow(sales_journal) * 100

cat("Sales Cutoff Errors:\n")
print(sales_cutoff_errors)

cat("\nNumber of Sales Cutoff Errors: ", cutoff_error_count, "\n")
cat("\nPercentage of Sales Cutoff Errors in the Total Sales Journal: ", cutoff_error_percentage, "%\n")

```


# Question 7: Check if AR balances are fairly stated

```{r}

library(readr)
library(pwr)
real_world_fyear_end_ar_ledger <- read.csv("/Users/abhinavram/Documents/Sem 2 UIC/IDS 523 Audit/auditanalytics_final_files/real_world_fyear_end_ar_ledger.csv")
sum(real_world_fyear_end_ar_ledger$amount)
nrow(real_world_fyear_end_ar_ledger)
total_ar_balance <- sum(real_world_fyear_end_ar_ledger$amount)
total_inovices <- nrow(real_world_fyear_end_ar_ledger)
average_ar_balance <- total_ar_balance/total_inovices
delta  <- 0.05 * average_ar_balance
sigma  <- 0.3 * average_ar_balance
effect <- delta/sigma
sample <- pwr.t.test(d = effect, sig.level = 0.05, power = 0.95, type = "one.sample", alternative = "greater")
estimated_error <- (sample$n * delta) / total_inovices

cat("\n attribute sample size for occurrence of error =", ceiling(sample$n), "\n estimated error in population = " , estimated_error,
    "\n Error is tolearable and, no need to extend the sample size",
    "\n AR balances are Fairly Stated")




```


# Question 8: Inventory Ommissions and Duplicates

```{r}

purchase_journal<-read.csv("/Users/abhinavram/Documents/Sem 2 UIC/IDS 523 Audit/auditanalytics_final_files/purchase_journal.csv")

receiver_journal<-read.csv("/Users/abhinavram/Documents/Sem 2 UIC/IDS 523 Audit/auditanalytics_final_files/receiver_journal.csv")

cat("Total number of purchase order numbers = ",nrow(purchase_journal))
cat("\nTotal number of receiver numbers = ",nrow(receiver_journal))

# Calculate percentage of duplicate purchase order numbers
dup_purchase <- purchase_journal[duplicated(purchase_journal$po_no), ]
perc_dup_purchase <- round(nrow(dup_purchase) / nrow(purchase_journal) * 100, 2)
cat("\nNumber of duplicate purchases = ", nrow(dup_purchase))
cat("\nPercentage of duplicate purchases = ", perc_dup_purchase)

# Calculate percentage of duplicate receiver numbers
dup_receiver <- receiver_journal[duplicated(receiver_journal$receiver_no), ]
perc_dup_receiver <- round(nrow(dup_receiver) / nrow(receiver_journal) * 100, 2)
cat("\nNumber of duplicate receiver_no in receiver_journal = ", nrow(dup_receiver))
cat("\nPercentage of duplicate receiver_no in receiver_journal = ", perc_dup_receiver)

# Calculate percentage of omitted purchase order numbers
po_numbers <- as.numeric(gsub("\\D", "", purchase_journal$po_no))
omitted_po <- setdiff(min(po_numbers):max(po_numbers), po_numbers)
perc_omitted_po <- round(length(omitted_po) / nrow(purchase_journal) * 100, 2)
cat("\nNumber of omitted purchase records = ", length(omitted_po))
cat("\nPercentage of omitted purchase records = ", perc_omitted_po)

# Calculate percentage of omitted receiver numbers
receiver_numbers <- as.numeric(gsub("\\D", "", receiver_journal$receiver_no))
omitted_receiver <- setdiff(min(receiver_numbers):max(receiver_numbers), receiver_numbers)
perc_omitted_receiver <- round(length(omitted_receiver) / nrow(receiver_journal) * 100, 2)
cat("\nNumber of omitted receiver records = ", length(omitted_receiver))
cat("\nPercentage of omitted receiver records = ", perc_omitted_receiver)

# Print results and determine if systems are in-control or out-of-control
tolerable_error_rate <- 1

if (perc_dup_receiver <= tolerable_error_rate && perc_omitted_receiver <= tolerable_error_rate) {
  cat("\nDuplicates in ReceiverNumbers: ", perc_dup_receiver, "% - Incontrol")
  cat("\nOmmissions in ReceiverNumbers: ", perc_omitted_receiver, "% - Incontrol")
} else {
  cat("\nDuplicates in ReceiverNumbers: ", perc_dup_receiver, "% - Out of control")
  cat("\nOmmissions in ReceiverNumbers: ", perc_omitted_receiver, "% - Out of control")
}

if (perc_dup_purchase > tolerable_error_rate || perc_omitted_po > tolerable_error_rate) {
  cat("\nDuplicates in PurchaseOrders: ", perc_dup_purchase, "% - Out of control")
  cat("\nOmmissions in PurchaseOrders: ", perc_omitted_po, "% - Out of control")
} else {
  cat("\nDuplicates in PurchaseOrders: ", perc_dup_purchase, "% - Incontrol")
  cat("\nOmmissions in PurchaseOrders: ", perc_omitted_po, "% - Incontrol")
}

```

# Question 9: Inventory Stock on Hand with a Turnover of Less than 5 times

```{r}
# Loading libraries
library(tidyverse)

perpetual_inventory_ledger <- read.csv("/Users/abhinavram/Documents/Sem 2 UIC/IDS 523 Audit/auditanalytics_final_files/perpetual_inventory_ledger.csv")

sales_journal <- read.csv("/Users/abhinavram/Documents/Sem 2 UIC/IDS 523 Audit/auditanalytics_final_files/sales_journal.csv")


# Calculating sales per SKU
sales_of_sku_per_year <- sales_journal %>%
  group_by(sku) %>%
  summarise(sales_rate = sum(sales_count, na.rm = TRUE))

# Getting the latest unit_cost for each SKU
unit_costs <- sales_journal %>%
  group_by(sku) %>%
  arrange(desc(invoice_date)) %>% # Arrange by date in descending order
  slice(1) %>% # Select the first row (the most recent one) for each SKU
  select(sku, unit_cost)

# Calculating the product of unit_cost and sales_rate
multiply <- sales_of_sku_per_year %>%
  left_join(unit_costs, by = "sku") %>%
  mutate(multiply1 = unit_cost * sales_rate)

# Calculating the stock on hand for each SKU
perpetual_inventory_ledger1 <- perpetual_inventory_ledger %>%
  group_by(sku) %>%
  summarise(stock = sum(stock_on_hand, na.rm = TRUE))

# Calculating the quantity of annual sales and filter for SKUs with less than 5 times turnover
quantityannualsales <- perpetual_inventory_ledger1 %>%
  left_join(multiply, by = "sku") %>%
  mutate(quantity_of_annual_sales = multiply1 / stock) %>%
  select(sku, quantity_of_annual_sales) %>%
  filter(quantity_of_annual_sales < 5)

# Displaying the result
quantityannualsales
```

# Question 10: Inventory Stock on Hand Lower of Cost or Market

```{r}
library(tidyverse)
real_world_ye_inventory[is.na(real_world_ye_inventory)] <- 0
nrv <- perpetual_inventory_ledger %>%
group_by(sku) %>%
slice(n()) %>% ## the final slice, by SKU, will be what is in-stock at year end
left_join(real_world_ye_inventory, by="sku") %>%
filter(exception != "No exception, count is accurate") %>%
mutate(err_perpetual = stock_on_hand - ye_stock_on_hand) %>%
select(sku,
stock_on_hand,
ye_stock_on_hand,
unit_cost,
actual_unit_market,
err_perpetual,
count_exception,
exception) %>%
mutate(ye_cost = stock_on_hand * unit_cost,
ye_market = ye_stock_on_hand * actual_unit_market,
sales_commission = ye_market * 0.1,
ye_market_with_commission = ye_market + sales_commission,
nrv = ye_market_with_commission - ye_cost) %>%
select(sku, ye_cost, ye_market, ye_market_with_commission , nrv) %>%
as.data.frame()
nrv

```


# Question 11: At year end the client counts the physical inventory and reconciles to Perpetual Inventory System


```{r }

perpetual_inventory_ledger<- read_csv("/Users/abhinavram/Documents/Sem 2 UIC/IDS 523 Audit/auditanalytics_final_files/perpetual_inventory_ledger.csv")


population_size=nrow(real_world_ye_inventory)

# Set the tolerable error amount
tolerable_error <- 10000000

# Calculate the minimum sample size required
z_value <- qnorm(0.975) # for 95% confidence level
p <- 0.5 # assuming a conservative estimate of 50% errors
e <- tolerable_error
min_sample_size <- ceiling((z_value^2 * p * (1 - p)) / (e^2))

N<- min_sample_size

sample_invoices <-  perpetual_inventory_ledger[sample(nrow(perpetual_inventory_ledger), N),]


inventory_count_differences <-
  perpetual_inventory_ledger %>% 
  group_by(sku) %>% 
  inner_join(real_world_ye_inventory, by="sku") %>% 
  slice(n()) %>%    # the final slice, by SKU, will be what is in-stock at year end
  filter(exception != "No exception, count is accurate") %>% 
  mutate(err_perpetual = stock_on_hand - ye_stock_on_hand) %>% 
  select(sku, err_perpetual, unit_cost, count_exception, exception, actual_unit_market) %>% 
  as.data.frame()
 
inventory_count_differences 

ratio<- population_size/N

error<- (sum(inventory_count_differences$err_perpetual))*ratio


cat("Estimated error in the population:", error, "\n")
cat("As error rate is less than tolerable rate we can conclude that inventory balance is materially fairly presented.")

```

# Estimated error in the population: 212000 

# Audit Opinion for Question 12 is written in the presentation.



